FISTA with adaptive discretisations

### 2021-01-15 paper.tex
Added Parpas reference
attempt to shrink text

@@

%In many variational approaches to inverse problems, optimisation problems are often written in a continuous setting but solved with a discrete optimisation algorithm. Consider what happens when $\sqrt{e^2}$ is computed in your favourite programming language. We have asked a `continuous' question and have been given a `discrete' answer with very well controlled error bounds.
%
%The parallel to this for optimisation is that we still do not have infinite computing memory or processing power, but we still want to compute minimisers to a known accuracy with efficient implementations. It is typically much harder to control discretisation errors in optimisation problems, for instance the $\Gamma$-convergence of discrete total variation \citep{Bartels2012,Bartels2015} or the approximation errors for wavelets/curvelets/sheerlets \citep{Mallat1999,Candes2004,Guo2007}. These are typically quite weak guarantees valid for asymptotic resolution/time and are hard to quantify in a specific example. In this work we propose an algorithm with three aims:
%\begin{itemize}
%	\item All computations are discrete but the asymptotic reconstruction is exact, even for infinite dimensional reconstructions in a Banach space, with a guaranteed rate.
%	\item Distance to the infinite dimensional minimiser can be quantified.
%	\item Discretisation is adaptively optimised for the particular minimiser.
%\end{itemize}
%
%We choose to modify the FISTA algorithm because it is quite general and performs very well in discrete optimisation problems. The strategy will be to follow the standard FISTA algorithm as closely as possible. The only difference is that at each iteration, computations will only be performed on a finite dimensional subspace. 
%
%One key issue is that FISTA is intrinsically designed to converge only in $L^2$. In finite dimensions there is no issue, all norms are equal and so convergence rates are equal for all problems up to scaling constants. On the other hand, in infinite dimensions there exist minimisers which are not in $L^2$. In this case, we find that the rate will always be slower than for standard FISTA but still at a guaranteed rate.
%
%We perform two numerical experiments where the minimiser is in $L^1\setminus L^2$ which demonstrates the reduced rate. We also show an example where the minimiser is in $\ell^2(\R^m)$ for some $m$ which is not known a priori and observe that the refining FISTA algorithm achieves linear convergence.

@@

The contribution of this work is a modified FISTA algorithm which achieves two key properties:
\begin{itemize}
	\item we extend the convergence results to the case $\varf0^*\in\bar{\F H}\setminus\F H$ where $\spcf0\coloneqq \bar{\F H}$ is a Banach space. The proven rate is slower than the optimal $n^{-2}$, but appears to be quite sharp in numerical examples.
	\item we allow for adaptive discretisation such that $\varf0_n\in\spcf0^n\subset\F H$. With simple guarantees on the rates of convergence, we show analytically that applying our FISTA algorithm with inexact discretisation does not reduce the proven rate of convergence. In the Hilbert space setting this is the original $n^{-2}$ rate. In numerical examples we demonstrate that greedy adaptive discretisation strategies can achieve much greater performance than uniform discretisation in both measures of time and computer memory requirements.
\end{itemize}
\todo{There is much overlap between the techniques used in this work and those used in the literature of inexact optimisation, although we believe that our interpretation for the use outside of Hilbert spaces is the key novelty.} Some other works for FISTA-like algorithms include \citep{Jiang2012,Villa2013}. Of particular note, our stability estimate for FISTA  in \Cref{thm: exponential FISTA convergence} is very similar to \cite[Proposition 2]{Schmidt2011} and \cite[Proposition 3.3]{Aujol2015}. This is then used to analyse the convergence properties in our more general Banach space setting where all sources of inexactness come from subspace approximations.

@@

Without loss of generality we will assume $L=1$, i.e. $\nabla\func1$ is 1-Lipschitz. To get the general statement of any of the results which follow, replace $\Func0$ with $\frac{\Func0}{L}$. Implicitly, the only impact of $L$ is scaling the convergence rates which are only given up to a constant factor anyway. A more important distinction is to say that the properties of $\func1$ and $\func2$ are stated with respect to the Hilbert norm. In particular,
$$\norm{\nabla\func1(\varf0)- \nabla\func1(\varf1)} \leq \norm{\varf0-\varf1}$$
for all $\varf0,\varf1\in\F H$ and $\func2$ is called `simple' if
$$\argmin_{\varf0\in\tilde{\spcf0}}\tfrac12\norm{\varf0-\varf1}^2 + \func2(\varf0) $$
is exactly computable for all $\varf1\in\F H$ and all $\tilde{\spcf0}\in\{\spcf0^n\}_{n=0}^\infty$.

@@

\begin{lemma}\label{thm: sufficiently fast}
	\Copy{thm: sufficiently fast}{
		Suppose $\spcf0^n, \varf0_n, \varf1_n$ and $n_k$ satisfy the conditions of \Cref{thm: mini exponential FISTA convergence} and $\{\tilde{\spcf0}^k\}$ forms an $(\vars3_{\Varf0},\vars3_{\Func0})$-discretisation for $\Func0$ with
		$$\tilde{\varf2}_k\in \left[\argmin_{\varf0\in\tilde{\spcf0}^k}\Func0(\varf0)\right]\cap\spcf0^{n_k}\cap\spcf0^{n_k+1}\cap\ldots\cap\spcf0^{n_{k+1}-1}.$$
		If either:
		\begin{itemize}
			\item $\vars3_{\Varf0}>1$ and $n_k^2\lesssim \vars3_{\Func0}^k\vars3_{\Varf0}^{2k}$,
			\item or $\vars3_{\Varf0}=1$, $\sum_{k=1}^\infty n_k^2\vars3_{\Func0}^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}<\infty,$
		\end{itemize}
		then
		$$ \Func0_0(\varf0_N) \lesssim \frac{\vars3_{\Varf0}^{2K}}{N^2}$$
		for all $n_K\leq N< n_{K+1}$.}
\end{lemma}

\noindent The proof is given in \cref{app:thm: sufficiently fast}. We make two observations of the optimality of \Cref{thm: sufficiently fast}:
\begin{itemize}
	\item The convergence guarantee for $N$ iterations of classical FISTA in the space $\spcf0^N$ is 
	$$ \Func0_0(\varf0_N) \lesssim \frac{\norm{\varf0_0-\argmin_{\varf0\in\spcf0^N}\Func0(\varf0)}^2}{N^2} + \min_{\varf0\in\spcf0^N}\Func0_0(\varf0) \lesssim \frac{\vars3_{\Varf0}^{2K}}{N^2} + \vars3_{\Func0}^{-K}.$$
	This is equivalent to \Cref{thm: sufficiently fast} under the conditions on $n_k$.
	\item If $\spcf0$ is finite dimensional then the $\vars3_{\Varf0}=1$ is almost trivially satisfied. Norms in finite dimensions are equivalent and any discretisation can be achieved with a finite number of refinements (i.e. the sums over $k$ are finite). 
\end{itemize}

@@

As a special case, note that if $\vars3_{\Varf0}=1$ then \Cref{thm: sufficiently fast} already gives the optimal $O(N^{-2})$ convergence rate. This is in fact a special case of \citet{Aujol2015}. If $\varf0^*\in\F H$ then it is not possible to refine `too quickly' and the following lemma is not needed.

@@

		\bigbreak\noindent Suppose that either:
		\begin{itemize}
			\item $\vars3_{\Varf0}>1$ and $n_k^2\simeq \vars3_{\Func0}^k\vars3_{\Varf0}^{2k}$,
			\item or $\vars3_{\Varf0}=1$, $\sum_{k=1}^\infty n_k^2\vars3_{\Func0}^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}<\infty$,
		\end{itemize}
		then
		$$\Func0_0(\varf0_N) \lesssim \frac{1}{N^{2(1-\kappa)}}\qquad \text{ where } \qquad \kappa = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2}$$
		uniformly for $N\in\F N$.
	}
\end{theorem}
\noindent This theorem is very easy to implement and requires very little knowledge of how to estimate $\Func0_0(\varf0_n)$. So long as $\vars3_{\Varf0}$ and $\vars3_{\Func0}$ can be computed analytically, choosing $\tilde{\spcf0}^k$ to be uniform discretisations and $\tilde{\spcf0}^k = \spcf0^{n_k}=\ldots= \spcf0^{n_{k+1}-1}$ will give the stated convergence rate.

@@

While \Cref{thm: exponential FISTA convergence} also allows for spatial adaptivity, we will simplify the requirements to a form which seems more practical for implementation and also allows for time adaptivity. We make the assumption that the choice of $n_k$ depends only on the value $\Func0(u^{n_k-1})$ and therefore make $u^{n_k-1}$ the root of our requirements. \cref{thm: sufficiently fast} suggests that a good refinement time strategy is to choose $n_k$ to be the minimal integer such that $\Func0_0(\varf0^{n_k-1})\lesssim \vars3_{\Func0}^{-k}$. However, the value of $\Func0_0$ may be hard to estimate and so we introduce a `backstop' condition which ensures that convergence is no slower than the rate given by \Cref{thm: exponential FISTA convergence}. The only constraint on spatial adaptivity is the piecewise monotone inclusion and that $\spcf0^{n_k} \ni u^{n_k-1}$. This is confirmed by the following theorem.
\begin{theorem}\label{thm: stronger exponential FISTA convergence}
	\Copy{thm: stronger exponential FISTA convergence}{
		Let $\{\spcf0^n\subset\F H\cap\spcf0\st n\in\F N\}$ be a sequence of subspaces and $n_k\in\F N$ a monotone increasing sequence such that
		$$ \tilde{\spcf0}^k \coloneqq \spcf0^{n_k}\ni \varf0_{n_k-1}, \qquad \tilde{\varf2}_k \in \left[\argmin_{\varf0\in\tilde{\spcf0}^k}\Func0(\varf0)\right]\cap\spcf0^{n_k+1} \cap\ldots\cap \spcf0^{n_{k+1}-1}$$
		for all $k\in\F N$.
		Compute $\varf0_n$, $\varf1_n$ by \Cref{alg: refining FISTA}.

@@

\noindent The proof is given in \cref{app:thm: practical refinement criteria}. The refinement criteria described by \Cref{thm: practical refinement criteria} can be split into two groups. Cases (1), (3), and (4) justify that any choice of $\tilde{\spcf0}^k\ni\varf0_k$ (or $\varf0_{n_k-1}$ in \Cref{thm: stronger exponential FISTA convergence}) satisfies the required conditions. In cases (2) and (5), $\varf0_k$ is sufficient to choose the refinement time $n_k$ but not the subspace $\tilde{\spcf0}^k$. In these cases one could choose $\tilde{\spcf0}^k$ to be a uniform discretisation with apriori estimates, for example.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{General examples}\label{sec: general examples}
In this work we consider the Lasso problem to be both a motivating example and source of numerical examples, however, \Cref{alg: refining FISTA} is much more broadly applicable. The aim of this section is to justify this claim by demonstrating that the constants $\vars3_{\Varf0}$ and $\vars3_{\Func0}$ can be easily computed in many examples.

@@

	We say that $(\tilde{\spcf0}^k)_k$ is \emph{of order $p$} if
	$$\min_{\varf2\in\tilde{\spcf0}^k}\Norm{\varf2-\varf0^*} \lesssim_{\varf0^*} \meshsize^{kp}.$$
	We allow the implicit constant to have any dependence on $\varf0^*$ so long as it is finite. For example, in the case of Sobolev spaces we would expect an inequality of the form
	$\min_{\varf2\in\tilde{\spcf0}^k}\norm{\varf2-\varf0^*}_{W^{0,q}} \lesssim \meshsize^{kp}\norm{\varf0^*}_{W^{p,q}}$ \citep{Strang1972}.
\end{definition}
We note that any piecewise polynomial finite element space can be used to form a $\meshsize$-refining sequence of subspaces. Wavelets almost satisfy this definition but not the support condition, each space contains every scale of wavelet. \todo{Is that true? If the wavelet has compact support then it should admit a basis with disjoint support?} Similarly, a Fourier basis does satisfy the scaling properties but each basis vector has global support. Both of these exceptions are important and could be accounted for with further analysis but we focus on the more standard finite element case. The following theorem shows that all finite element spaces achieve the same basic rates of convergence in \Cref{alg: refining FISTA}, depending on the particular properties of $\Norm\cdot$ and $\Func0$.
\begin{theorem}\label{thm: generic a_U and a_E}
	Suppose $\F H= L^2(\Domain)$ for some compact domain $\Domain\subset\R^d$ and $\norm\cdot_q\lesssim \Norm\cdot$ for some $q\in[1,\infty]$. If $(\tilde{\spcf0}^k)_k$ is a sequence of $\meshsize$-refining finite element spaces of order $p$ then
	\begin{align*}
		\vars3_{\Varf0} &\leq \splitln{1}{q\geq 2}{\sqrt{\meshsize^{-d}}}{q<2},
		& \vars3_{\Func0} &\geq \begin{cases}
			\meshsize^{-p} & \Func0 \text{ is $\Norm\cdot$-Lipschitz at }\varf0^*
			\\\meshsize^{-2p} & \Func0 \text{ is $\Norm\cdot$-smooth at }\varf0^*
		\end{cases}\ .
	\end{align*}
	If $\func2$ is not Lipschitz at $\varf0^*$ but $\func1$ is $\Norm\cdot$-Lipschitz and 
	$$\min_{\varf2\in\tilde{\spcf0}^k}\left\{\Norm{\varf2-\varf0^*}\st \func2(\varf2)\leq \func2(\varf0^*)\right\} \lesssim \min_{\varf2\in\tilde{\spcf0}^k}\Norm{\varf2-\varf0^*},$$
	then, again, $\vars3_{\Func0}\geq\meshsize^{-p}$. 
\end{theorem}
The proof of this theorem is in \Cref{app: generic a_U and a_E}. The main take-home for this theorem is that the computation of $\vars3_{\Varf0}$ and $\vars3_{\Func0}$ is typically very simple and clear given a particular choice of $\Norm\cdot$ and $\Func0$. The only non-trivial case is when $\func2$ is not smooth enough to estimate directly. So long as $\tilde{\spcf0}^k$ are sufficiently dense in sublevel sets of $\func2$, the smoothness of $\func1$ is sufficient to give a convergence rate.

@@

Application of \Cref{thm: generic a_U and a_E} gives $\vars3_{\Varf0}=2^{\frac d2}$ although although $\norm\cdot_1$ is too strong a metric and results in $\vars3_{\Func0}=2^0$. This can be seen because for any $\varf0\in L^1(\Domain^d)$, 
\begin{equation}
	\norm{\varf0-\delta_0}_1 = \sup_{\vard0\in C(\Domain^d), \norm{\vard0}_\infty\leq 1} \IP{\vard0}{\varf0-\delta_0} = \norm{\varf0}_1+\norm{\delta_0}_1 \geq 1.
\end{equation}
To improve our estimate of $\vars3_{\Func0}$ requires additional assumptions on $\A$. For any $\varf2\in\tilde{\spcf0}^k$ such that $\norm{\varf2}_1\leq \norm{\varf0^*}_1$, we compute
\begin{align}
	\Func0(\varf2)-\Func0(\varf0^*) &=  \tfrac12\norm{\A\varf2-\vdata0}^2 - \tfrac12\norm{\A\varf0^*-\vdata0}^2 + \mu\left(\Norm{\varf2}-\Norm{\varf0^*}\right)
	\\&\leq \tfrac12\norm{\A\varf2-\vdata0}^2 - \tfrac12\norm{\A\varf0^*-\vdata0}^2
	\\&= \IP{\tfrac12\A(\varf2+\varf0^*)-\vdata0}{\A(\varf2-\varf0^*)}
	\\&\leq \max(\norm{\A\varf2-\vdata0},\norm{\A\varf0^*-\vdata0})\norm{\A(\varf2-\varf0^*)} 
	\\&\leq \sqrt{2\Func0(\varf2)}\norm{\A(\varf2-\varf0^*)}.
\end{align}
Choosing $\varf2 = \tilde\Pi_k\varf0^*$ for any bounded linear extension to the orthogonal projection $\tilde\Pi_k\colon\spcf0\to\tilde{\spcf0}^k$ gives
$$\Func0(\varf2)-\Func0(\varf0^*) \lesssim \norm{\A(\tilde\Pi_k\varf0^*-\varf0^*)} \leq \norm{(\tilde\Pi_k-\id)\A^*}_{\ell^2\to L^\infty}\Norm{\varf0^*}.$$
Expanding the operator norm further,
\begin{multline}
	\norm{(\tilde\Pi_k-\id)\A^*}_{\ell^2\to L^\infty} = \sup_{\vvarx3\in\R^m}\max_{\vvarx0\in\Domain} \frac{|[\tilde\Pi_k\A^*\vvarx3](\vvarx0)-[\A^*\vvarx3](\vvarx0)|}{\norm{\vvarx3}_{\ell^2}} \leq \sup_{\vvarx3\in\R^m} \frac{\sqrt{d}2^{-k}\norm{\nabla[\A^*\vvarx3]}_{L^\infty}}{\norm{\vvarx3}_{\ell^2}} 
	\\= \sqrt{d}2^{-k}|\A^*|_{\ell^2\to C^1}
\end{multline}
where $\sqrt d2^{-k}$ is the diameter of a $d$ dimensional pixel of side length $2^{-k}$.

@@

Inserting the computed rates into \Cref{thm: exponential FISTA convergence} or \Cref{thm: stronger exponential FISTA convergence} gives the guaranteed convergence rate
\begin{equation}\label{eq: Lasso energy rate}
	\kappa = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2} = \frac{d}{1+d} \quad \implies \quad \Func0_0(\varf0_n) \lesssim n^{-2(1-\kappa)} = n^{-\frac{2}{1+d}}.
\end{equation}
This rate can be used to infer the required resolution at each iteration, in particular on iteration $n$ with $n^2\sim (\vars3_{\Func0}\vars3_{\Varf0}^2)^k$ we expect the resolution to be
\begin{equation}\label{eq: Lasso resolution rate}
	2^{-k} = \left(\vars3_{\Func0}\vars3_{\Varf0}^2\right)^{\frac{k}{1+d}} \sim n^{-\frac{2}{1+d}}.
\end{equation}

>>>>

%\renewcommand{\Paste}[1]{}

@@

@@

The contribution of this work is a modified FISTA algorithm which achieves two key properties:
\begin{itemize}
	\item we extend the convergence results to the case $\varf0^*\in\bar{\F H}\setminus\F H$ where $\spcf0\coloneqq \bar{\F H}$ is a Banach space. The proven rate is slower than the optimal $n^{-2}$, but appears to be quite sharp in numerical examples.
	\item we allow for adaptive discretisation such that $\varf0_n\in\spcf0^n\subset\F H$. With simple guarantees on the rates of convergence, we show analytically that applying our FISTA algorithm with inexact discretisation does not reduce the proven rate of convergence. In the Hilbert space setting this is the original $n^{-2}$ rate. In numerical examples we demonstrate that greedy adaptive discretisation strategies can achieve much greater performance than uniform discretisation in both measures of time and computer memory requirements.
\end{itemize}
\todo{There is much overlap between the techniques used in this work and those used in the literature of inexact optimisation, although we believe that our interpretation for the use outside of Hilbert spaces is the key novelty.} Some other works for FISTA-like algorithms include \citep{Jiang2012,Villa2013}. Of particular note, our stability estimate for FISTA  in \Cref{thm: exponential FISTA convergence} is very similar to \cite[Proposition 2]{Schmidt2011} and \cite[Proposition 3.3]{Aujol2015}. This is then used to analyse the convergence properties in our more general Banach space setting where all sources of inexactness come from subspace approximations. The ideas of \citet{Parpas2017} are similar although in application to the proximal gradient method with an additional smoothing on the functional $\func2$. The permitted refinement steps are also more broad in our work.

@@

Without loss of generality we will assume $L=1$, i.e. $\nabla\func1$ is 1-Lipschitz. To get the general statement of any of the results which follow, replace $\Func0$ with $\frac{\Func0}{L}$. A more important distinction is to say that the properties of $\func1$ and $\func2$ are stated with respect to the Hilbert norm. In particular,
\begin{equation}
	\norm{\nabla\func1(\varf0)- \nabla\func1(\varf1)} \leq \norm{\varf0-\varf1}
\end{equation}
for all $\varf0,\varf1\in\F H$ and $\func2$ is called `simple' if
\begin{equation}
	\argmin_{\varf0\in\tilde{\spcf0}}\tfrac12\norm{\varf0-\varf1}^2 + \func2(\varf0)
\end{equation}
is exactly computable for all $\varf1\in\F H$ and all $\tilde{\spcf0}\in\{\spcf0^n\}_{n=0}^\infty$.

@@

\begin{lemma}\label{thm: sufficiently fast}
	\Copy{thm: sufficiently fast}{
		Suppose $\spcf0^n, \varf0_n, \varf1_n$ and $n_k$ satisfy the conditions of \Cref{thm: mini exponential FISTA convergence} and $\{\tilde{\spcf0}^k\}$ forms an $(\vars3_{\Varf0},\vars3_{\Func0})$-discretisation for $\Func0$ with
		$$\tilde{\varf2}_k\in \left[\argmin_{\varf0\in\tilde{\spcf0}^k}\Func0(\varf0)\right]\cap\spcf0^{n_k}\cap\spcf0^{n_k+1}\cap\ldots\cap\spcf0^{n_{k+1}-1}.$$
		If either:
		\begin{itemize}
			\item $\vars3_{\Varf0}>1$ and $n_k^2\lesssim \vars3_{\Func0}^k\vars3_{\Varf0}^{2k}$,
			\item or $\vars3_{\Varf0}=1$, $\sum_{k=1}^\infty n_k^2\vars3_{\Func0}^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}<\infty,$
		\end{itemize}
		then
		$$ \Func0_0(\varf0_N) \lesssim \frac{\vars3_{\Varf0}^{2K}}{N^2}\qquad \text{for all}\qquad n_K\leq N< n_{K+1}.$$
		}
\end{lemma}

\noindent The proof is given in \cref{app:thm: sufficiently fast}. We make two observations of the optimality of \Cref{thm: sufficiently fast}:
\begin{itemize}
	\item The convergence guarantee for $N$ iterations of classical FISTA in the space $\spcf0^N$ is 
	$$ \Func0_0(\varf0_N) \lesssim \frac{\norm{\varf0_0-\argmin_{\varf0\in\spcf0^N}\Func0(\varf0)}^2}{N^2} + \min_{\varf0\in\spcf0^N}\Func0_0(\varf0) \lesssim \frac{\vars3_{\Varf0}^{2K}}{N^2} + \vars3_{\Func0}^{-K}.$$
	This is equivalent to \Cref{thm: sufficiently fast} after the assumptions on $n_k$.
	\item If $\spcf0$ is finite dimensional then the $\vars3_{\Varf0}=1$ is almost trivially satisfied. Norms in finite dimensions are equivalent and any discretisation can be achieved with a finite number of refinements (i.e. the sums over $k$ are finite). 
\end{itemize}

@@

As a special case, note that if $\vars3_{\Varf0}=1$ then \Cref{thm: sufficiently fast} already gives the optimal $O(N^{-2})$ convergence rate. This is in fact a special case of \citet{Aujol2015}. If $\varf0^*\in\F H$, then it is not possible to refine `too quickly' and the following lemma is not needed.

@@

		\bigbreak\noindent Suppose that either:
		\begin{itemize}
			\item $\vars3_{\Varf0}>1$ and $n_k^2\simeq \vars3_{\Func0}^k\vars3_{\Varf0}^{2k}$,
			\item or $\vars3_{\Varf0}=1$, $\sum_{k=1}^\infty n_k^2\vars3_{\Func0}^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}<\infty$,
		\end{itemize}
		then
		$$\Func0_0(\varf0_N) \lesssim \frac{1}{N^{2(1-\kappa)}}\qquad \text{ where } \qquad \kappa = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2} \qquad \text{uniformly for } N\in\F N.$$
		}
\end{theorem}
\noindent This theorem is very easy to implement numerically and requires very little knowledge of how to estimate $\Func0_0(\varf0_n)$. So long as $\vars3_{\Varf0}$ and $\vars3_{\Func0}$ can be computed analytically, choosing $\tilde{\spcf0}^k$ to be uniform discretisations and $\tilde{\spcf0}^k = \spcf0^{n_k}=\ldots= \spcf0^{n_{k+1}-1}$ will give the stated convergence rate.

@@

While \Cref{thm: exponential FISTA convergence} also allows for spatial adaptivity, we will simplify the requirements to a form which seems more practical for implementation and also allows for time adaptivity. We make the assumption that the choice of $n_k$ depends only on the value $\Func0(u^{n_k-1})$ and therefore make $u^{n_k-1}$ the root of our requirements. \cref{thm: sufficiently fast} suggests that a good refinement time strategy is to choose $n_k$ to be the minimal integer such that $\Func0_0(\varf0^{n_k-1})\lesssim \vars3_{\Func0}^{-k}$. However, the value of $\Func0_0$ may be hard to estimate and so we introduce a `backstop' condition which guarantees that convergence is no slower than the rate given by \Cref{thm: exponential FISTA convergence}. The only constraint on spatial adaptivity is the piecewise monotone inclusion and that $\spcf0^{n_k} \ni u^{n_k-1}$. This is confirmed by the following theorem.
\begin{theorem}\label{thm: stronger exponential FISTA convergence}
	\Copy{thm: stronger exponential FISTA convergence}{
		Let $\{\spcf0^n\subset\F H\cap\spcf0\st n\in\F N\}$ be a sequence of subspaces and $n_k\in\F N$ a monotone increasing sequence such that
		$$ \tilde{\spcf0}^k \coloneqq \spcf0^{n_k}\ni \varf0_{n_k-1}, \qquad \tilde{\varf2}_k \in \left[\argmin_{\varf0\in\tilde{\spcf0}^k}\Func0(\varf0)\right]\cap\spcf0^{n_k+1} \cap\ldots\cap \spcf0^{n_{k+1}-1}$$
		for all $k\in\F N$.
		Compute $\varf0_n$, $\varf1_n$ by \Cref{alg: refining FISTA}.

@@

\noindent The proof is given in \cref{app:thm: practical refinement criteria}. The refinement criteria described by \Cref{thm: practical refinement criteria} can be split into two groups. Cases (1), (3), and (4) justify that any choice of $\tilde{\spcf0}^k\ni\varf0_k$ (or $\varf0_{n_k-1}$ in \Cref{thm: stronger exponential FISTA convergence}) satisfies the required conditions. In cases (2) and (5), $\varf0_k$ is sufficient to choose the refinement time $n_k$ but not the subspace $\tilde{\spcf0}^k$. In these cases one could, for example, choose $\tilde{\spcf0}^k$ to be a uniform discretisation with apriori estimates.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{General examples}\label{sec: general examples}
In this work we consider the Lasso problem to be both a motivating example and source of numerical examples, however, \Cref{alg: refining FISTA} is much more broadly applicable. Here we justify this claim by demonstrating that the constants $\vars3_{\Varf0}$ and $\vars3_{\Func0}$ can be easily computed in many examples.

@@

	We say that $(\tilde{\spcf0}^k)_k$ is \emph{of order $p$} if
	$$\min_{\varf2\in\tilde{\spcf0}^k}\Norm{\varf2-\varf0^*} \lesssim_{\varf0^*} \meshsize^{kp}.$$
	We allow the implicit constant to have any dependence on $\varf0^*$ so long as it is finite. For example, in the case of Sobolev spaces we would expect an inequality of the form
	$\min_{\varf2\in\tilde{\spcf0}^k}\norm{\varf2-\varf0^*}_{W^{0,q}} \lesssim \meshsize^{kp}\norm{\varf0^*}_{W^{p,q}}$ \citep{Strang1972}.
\end{definition}
We note that any piecewise polynomial finite element space can be used to form a $\meshsize$-refining sequence of subspaces. Wavelets with a compactly supported basis satisfy this definition but it is hard to specify the value $N$ in general. Similarly, a Fourier basis does satisfy the scaling properties but each basis vector has global support. Both of these exceptions are important and could be accounted for with further analysis but we focus on the more standard finite element case. The following theorem shows that all finite element spaces achieve the same basic rates of convergence in \Cref{alg: refining FISTA}, depending on the particular properties of $\Norm\cdot$ and $\Func0$.
\begin{theorem}\label{thm: generic a_U and a_E}
	Suppose $\F H= L^2(\Domain)$ for some compact domain $\Domain\subset\R^d$ and $\norm\cdot_q\lesssim \Norm\cdot$ for some $q\in[1,\infty]$. If $(\tilde{\spcf0}^k)_k$ is a sequence of $\meshsize$-refining finite element spaces of order $p$ then
	\begin{align*}
		\vars3_{\Varf0} &\leq \splitln{1}{q\geq 2}{\sqrt{\meshsize^{-d}}}{q<2},
		& \vars3_{\Func0} &\geq \begin{cases}
			\meshsize^{-p} & \Func0 \text{ is $\Norm\cdot$-Lipschitz at }\varf0^*
			\\\meshsize^{-2p} & \Func0 \text{ is $\Norm\cdot$-smooth at }\varf0^*
		\end{cases}\ .
	\end{align*}
	If $\func2$ is not Lipschitz at $\varf0^*$ but $\func1$ is $\Norm\cdot$-Lipschitz and 
	$$\min_{\varf2\in\tilde{\spcf0}^k}\left\{\Norm{\varf2-\varf0^*}\st \func2(\varf2)\leq \func2(\varf0^*)\right\} \lesssim \min_{\varf2\in\tilde{\spcf0}^k}\Norm{\varf2-\varf0^*},$$
	then, again, $\vars3_{\Func0}\geq\meshsize^{-p}$. 
\end{theorem}
The proof of this theorem is in \Cref{app: generic a_U and a_E}. The main take-home for this theorem is that the computation of $\vars3_{\Varf0}$ and $\vars3_{\Func0}$ is typically very simple and clear given a particular choice of $\Norm\cdot$ and $\Func0$. The only non-trivial case is when $\func2$ is not smooth enough to estimate directly. So long as $\tilde{\spcf0}^k$ are sufficiently dense in sublevel sets of $\func2$, the smoothness of $\func1$ is sufficient to give a convergence rate.

@@

Application of \Cref{thm: generic a_U and a_E} gives $\vars3_{\Varf0}=2^{\frac d2}$ although although $\norm\cdot_1$ is too strong a metric and results in $\vars3_{\Func0}=2^0$. This can be seen because for any $\varf0\in L^1(\Domain^d)$, 
\begin{equation}
	\norm{\varf0-\delta_0}_1 = \sup_{\vard0\in C(\Domain^d), \norm{\vard0}_\infty\leq 1} \IP{\vard0}{\varf0-\delta_0} = \norm{\varf0}_1+\norm{\delta_0}_1 \geq 1.
\end{equation}
To improve our estimate of $\vars3_{\Func0}$ requires additional assumptions on $\A$. For any $\varf2\in\tilde{\spcf0}^k$ such that $\norm{\varf2}_1\leq \norm{\varf0^*}_1$, we compute
\begin{align}
	\Func0(\varf2)-\Func0(\varf0^*) &=  \tfrac12\norm{\A\varf2-\vdata0}^2 - \tfrac12\norm{\A\varf0^*-\vdata0}^2 + \mu\left(\Norm{\varf2}-\Norm{\varf0^*}\right)
	\\&\leq \tfrac12\norm{\A\varf2-\vdata0}^2 - \tfrac12\norm{\A\varf0^*-\vdata0}^2
	\\&= \IP{\tfrac12\A(\varf2+\varf0^*)-\vdata0}{\A(\varf2-\varf0^*)}
	\\&\leq \max(\norm{\A\varf2-\vdata0},\norm{\A\varf0^*-\vdata0})\norm{\A(\varf2-\varf0^*)} 
	\\&\leq \sqrt{2\Func0(\varf2)}\norm{\A(\varf2-\varf0^*)}.
\end{align}
Choosing $\varf2 = \tilde\Pi_k\varf0^*$ for any bounded linear extension to the orthogonal projection $\tilde\Pi_k\colon\spcf0\to\tilde{\spcf0}^k$ gives
$$\Func0(\varf2)-\Func0(\varf0^*) \lesssim \norm{\A(\tilde\Pi_k\varf0^*-\varf0^*)} \leq \norm{(\tilde\Pi_k-\id)\A^*}_{\ell^2\to L^\infty}\Norm{\varf0^*}.$$
Expanding the operator norm further,
\begin{equation}
	\norm{(\tilde\Pi_k-\id)\A^*}_{\ell^2\to L^\infty} = \sup_{\vvarx3\in\R^m}\max_{\vvarx0\in\Domain} \frac{|[\tilde\Pi_k\A^*\vvarx3](\vvarx0)-[\A^*\vvarx3](\vvarx0)|}{\norm{\vvarx3}_{\ell^2}} \leq \sup_{\vvarx3\in\R^m} \frac{\sqrt{d}2^{-k}\norm{\nabla[\A^*\vvarx3]}_{L^\infty}}{\norm{\vvarx3}_{\ell^2}} 
	= \sqrt{d}2^{-k}|\A^*|_{\ell^2\to C^1}
\end{equation}
where $\sqrt d2^{-k}$ is the diameter of a $d$ dimensional pixel of side length $2^{-k}$.

@@

Inserting the computed rates into \Cref{thm: exponential FISTA convergence} or \Cref{thm: stronger exponential FISTA convergence} gives the guaranteed convergence rate
\begin{equation}\label{eq: Lasso energy rate}
	\kappa = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2} = \frac{d}{1+d} \quad \implies \quad \Func0_0(\varf0_n) \lesssim n^{-2(1-\kappa)} = n^{-\frac{2}{1+d}}.
\end{equation}
This rate can be used to infer the required resolution at each iteration, in particular on iteration $n$ with $n^2\simeq (\vars3_{\Func0}\vars3_{\Varf0}^2)^k$ we expect the resolution to be
\begin{equation}\label{eq: Lasso resolution rate}
	2^{-k} = \left(\vars3_{\Func0}\vars3_{\Varf0}^2\right)^{\frac{k}{1+d}} \simeq n^{-\frac{2}{1+d}}.
\end{equation}
